Synchronizing Distributed Openflow Controllers\\{\Large Toward More Powerful In-Band Synchronization Objects

### 2015-03-08 inband-STN.tex
stefan iteration

%\title{Solving Consensus with OpenFlow}

@@

%\title{One-Shot Coordination of Distributed SDN Controllers:\\On the Consensus Power of OpenFlow}

%\title{Coordinating Distributed SDN Controllers:\\On the Consensus Power of OpenFlow}

%\title{On the Synchronization Power of OpenFlow}

\title{Coordinating Distributed SDN Controllers\\{\Large On the Synchronization Power of OpenFlow}}

@@

\begin{abstract}
Control planes of modern Software-Defined Networks (SDNs)
are \emph{distributed systems}: to ensure availability and fault-tolerance,
to improve load-balancing, and to reduce overheads, 
control modules are physically distributed.
%a redundant control plane
%can ensure a high availability, and a spatially distributed
%control plane allows to handle time-critical dataplane
%events close to their origins.
%[[PK this is a contraversial point, not important for us
%Moreover, an SDN may be managed by multiple administrators simultaneously.
%]]
However, a distributed control plane also introduces
new challenges, and in order to guarantee a consistent network operation,
the actions of different controllers may need to be synchronized and coordinated.
%[[PK sounds contradicting to the fact that such systems already exist
%However, not much is known today about how to implement
%such a distributed control plane.
%]]
In this paper, we show that powerful synchronization mechanisms
can be implemented \emph{in band}, simply using today's Openflow switches.
We show how to leverage standard OpenFlow features for providing simple
but general-purpose transactions with all-or-nothing semantics.
We then show that these transactions can be used to reach consensus
among software controllers and, what is more, to resolve
general conflicts between concurrent policy updates in an
effectively atomic way and \emph{on-the-fly}, without affecting the ongoing traffic.
%[[PK
%In particular, we describe the implementation of test\&set primitives
%as well as multi-writer consensus objects which can be used
%to implement atomic transactions with all-or-nothing semantics.
%Interestingly, our approach allows controllers to detect and
%resolve very general notions of conflicts, not only, e.g., conflicts
%due to overlapping flow spaces but also conflicts based on network load.
%]]
\end{abstract}

@@

By consolidating and outsourcing the control over the dataplane switches to a logically
centralized controller, the concept Software-Defined Networking (SDN)
simplifies network management and facilitates faster innovations,
as the (software) control plane is decoupled from the (hardware) data
plane and can evolve independently.
Moreover, Openflow, the standard SDN protocol today, introduces many generalizations,
in terms of traffic engineering, definition of flows, as well as in-band network functionalities,
by relying on a simple match-action paradigm which allows us to define
forwarding rules based not only on Layer-2, but also Layer-3 and Layer-4 (and possibly beyond)

While the logically centralized perspective offered by SDN is attractive,
there is a wide consensus that
the control plane should be physically \emph{distributed}. 
First, in order to provide a high availability as well as a minimal degree of
fault-tolerance, controllers should be redundant~\cite{onix,stn,onos}: a failure
of one controller can be masked by other controllers. Second, it has also been proposed
to distribute controllers \emph{spatially}, in order to handle latency-sensitive and
communication intensive control plane events close to their origin.~\cite{devoflow,kandoo,jukka,disco}
Third, larger SDNs are likely to be operated by multiple administrators or may even offer
participatory interfaces where different users can install and trigger policy changes
concurrently~\cite{participatory,stn}.
All this, however, comes with the need for \emph{synchronization}
and \emph{coordination} among the distributed controllers:
the network should be manipulated \emph{consistently}~\cite{cpc}.
This task is nontrivial~\cite{cap-theorem}.

@@

Consider, for example, the practical problem of
consistent installation of new forwarding policies, stipulating routes
that packets of different header spaces should follow across the
network~\cite{network-update,roger-hotnets,correct,stn}.
Installing conflicting forwarding rules, e.g., rules of the same priority defined over non-disjoint
flow spaces may lead to pathological network behavior (loops,
blackholes, etc.).
Similarly, installing diverging load-balancing policies may, 
when combined, \emph{increase} the load~\cite{log-cent}.
To render things more difficult, controllers may also fail, 
even before their updates have been completed.

%\textbf{} 
\paragraph{Our Contributions}
In this paper, we show that 
OpenFlow~\cite{of-spec}---the standard SDN protocol today---can
be used to provide strong controller synchronization and coordination 
primitives \emph{in-band}.
In particular, we show how to exploit the \emph{bundling}
feature and the $\textsf{OFPFF\_CHECK\_OVERLAP}$ flag,
to implement a \emph{shared memory} abstraction of the switch
and its flow tables: 
This shared memory 
allows us to implement
\emph{transactions} which may
atomically install multiple rules if certain compatibility conditions
are met, and which also facilitate \emph{packet-atomic} rule updates. 
These transactions can be \emph{multi-write} or even \emph{multi-modify}:
while not supporting
read operations, the transaction can be conditioned on the presence of
certain rules.
This ability is reminiscent of the hypothetical multi-write
shared-memory model which is known to enable solutions for important
synchronization problems, such as consensus, impossible to solve
otherwise~\cite{multi-objects}.
Indeed, we describe an algorithm that solves consensus among any
number of controllers using a single OpenFlow switch that accepts
bundle control messages.

Our results highlight the synchronization power hidden in the
OpenFlow standard. The synchronization mechanisms we describe in this
paper turn out to be crucial for ensuring consistency of network
management on the control plane while preserving consistency of the
traffic processed on the data plane.
As a case study, we also show how to replace the read-modify-write
primitives voluntarily assumed in the recent implementation of a
consistent composition system for forwarding policies~\cite{cpc}.
We also show how to use these transactions directly to implement a lightweight
forwarding policy composition system that requires no additional
synchronization on the control plane.
Finally, we explore how more general classes of policies can be composed
\emph{on-the-switch}.

@@

\paragraph{Organization}
The rest of the paper is organized as follows.
In Section~\ref{sec:model}, we overview the basics of the  OpenFlow
protocol and recall the bundle and \textsf{ofpff\_check\_overlap} features.
In Section~\ref{sec:sync}, we specify the implementations of our OpenFlow-based
sycnhronization mechanisms.
In Section~\ref{sec:application}, we discuss a few policy composition
algorithms that benefit from these mechanisms.
In Section~\ref{sec:conclusion}, we discuss open questions and related work.

@@

\section{OpenFlow Background}\label{sec:model}

The synchronization and coordination mechanisms presented in this paper
are based on Openflow~\cite{of-spec}, the standard SDN protocol today. In this section, 
we provide the necessary background.

es a small partial deployment of SDN and
out-sources the control over a set of Openflow switches to a logically centralized software controller.
Openflow, the standard SDN protocol today, is based on a match-action paradigm: the controller
can install simple rules on the Openflow switches.

In a nutshell, 
OpenFlow
is based on a match-action concept: OpenFlow switches store
rules (installed by the controller) consisting of a match and an
action part. A packet matched by a certain rule will be subject
to the associated action (to be immediately applied now or later). 
For example, an action can define a port to which the
matched packet should be forwarded, or add or change a tag
(a certain part in the packet header).
Each OpenFlow switch stores one or multiple flow tables,
each of which contains a set of rules (flow entries). Flow
tables form a pipeline, and flow entries are ordered according
to priorities: A packet arriving at a switch is first checked by
the rule of the highest priority in table 0: the fields of the
data packet are compared with the match fields of that rule,
and if they fit, some instructions (the actions) are executed;
subsequently, lower priority rules are checked. Depending on
the outcome of the table 0 processing, the packet may be sent
to additional flow tables in the pipeline; during the traversal of
the flow tables, the packet can be equipped with meta-tags to
transfer temporary information. Concretely, instructions can be
used to define additional tables to be visited (goto instruction),
to modify the set of to-be-applied actions (either by appending,
deleting, or modifying actions), or immediately apply some
actions to the packet. A meta-data field can be used to exchange
information between tables; it is part of the header that can be
inserted or removed from a packet via push and pop actions.
In general, a packet can be matched against any of its header
fields, and fields can be wildcarded and sometimes bitmasked.
(For instance, the meta-data field is maskable.) If no rule
matches, the packet is dropped.

\textbf{TODO LIRON: what else is missing? describe cookies?}

The Openflow standard (as of \textbf{TODO}) offers two interesting
features which we will exploit to implement the transactional memory abstraction:

@@

\item \textbf{Conditional Application:}
The Openflow standard also supports the \emph{conditional} application of
certain types of control messages. A specific $\textsf{OFPFF\_CHECK\_OVERLAP}$ flag
set in the control message containing a new forwarding rule ensures
that the rule is applied if and only if it does not \emph{conflict}
with any other rule currently in the flow table.
This in-band conflict detection mechanism, combined with bundling,
allows us to implement specific types of \emph{transactions} that may
atomically install multiple rules if certain compatibility conditions
are met.
\end{enumerate}

\section{Vision and Problem Statement}\label{sec:vision}

This paper shows how to implement a transactional memory abstraction using
standard Openflow switches. The shared memory abstraction of the 
flow table at the switch allows multiple controllers to atomically and 
concurrently read and modify state,
using the bundling feature to implement \emph{multi-write} or, more
widely, \emph{multi-modify} transactions. .
This ability allows us to solve consensus among any
number of controllers, using a single OpenFlow switch that accepts
bundle control messages.

These synchronization mechanisms can be used to
achieve consistent network management.
We first show that they can be used to replace the read-modify-write
primitives voluntarily assumed in the recent implementation of a
recent proposal for a consistent composition system for forwarding policies~\cite{cpc}.
We then use these transactions directly to implement a lightweight
forwarding policy composition system that requires no additional
synchronization on the control plane.
We then explore how more general classes of policies can be composed
\emph{on-the-switch}.

\section{Transactions in OpenFlow}\label{sec:sync}

@@

\begin{algorithm}[t]
    \caption{Advanced Update Algorithm}
    \label{alg:template}
    \begin{algorithmic}[1]
    \Require update requirements $U$, multi-set (with optional add validation) NEXT, set (with full validation) CUR.
    \Ensure installed policy is consistent with previous one
 		%\Do
 		\Repeat
 			\State read switch configuration
 			\State $my\_id\gets$ unused id (in both CUR and NEXT)
 			\State $pid\gets$ current policy id (only id in CUR)
 			\State NEXT.Add($pid$)
 			\State read switch configuration
 			\If {$pid\notin CUR$}
 				\State {\bf continue} (restart loop)
 			\EndIf
 			\State $cur\_policy\gets$ current policy
 			\State $new\_policy\gets apply\_update(cur\_policy,U)$
 			\startTransaction
 				\State NEXT.Add($my\_id$, True)
 				\State NEXT.Delete($my\_id$, True)
	 			\State DeleteGroup($pid$)
	 			\State AddGroup($my\_id$)
				\State NEXT.Delete($pid$)
	 			\State add $new\_policy$ rules %conditioned on $metadata=my\_a$
	 			%\State update policy selector to $my\_a$
 			\endTransaction
     		%\State $res \gets $ commit transaction
     	\Until{$res=True$}
     	%\doWhile{$res=False$}
%   \vspace{3mm}
			\Return

\begin{algorithm}[t]
    \caption{multi-set Add abstraction}
    \label{alg:template}
    \begin{algorithmic}[1]
    \Require id $i$, unique constrain (defaults to False) $b\_unique$, caller id $caller$ (defaults to self).
%    \Ensure
    		\State $value \gets i$
    		\State $mask \gets all-ones$
    		\State $cookie \gets i.caller$
    		\State $action \gets i.caller$
    		\If {$b\_unique$}
    			\State $flag \gets 0$
    		\Else
    		    \State $flag \gets CHECK\_OVERLAP$
    		\EndIf
    					\State $cmd\gets FlowMod(match=(value,mask), operation = ADD, cookie, flag, action) $
			\Return cmd
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[t]
    \caption{multi-set Delete abstraction}
    \label{alg:template}
    \begin{algorithmic}[1]
    \Require id $i$, unique constrain (defaults to False) $b\_unique$, caller id $caller$.
%    \Ensure
    		\State $value \gets C.i$
    		\State $mask \gets all-ones$
    		\State $cookie \gets C.i.caller$
    		\If {$b\_unique$}
    			\State $cmd\gets FlowMod(match=(value,mask), operation=DELETE) $
    		\Else
    		    \State $cmd\gets FlowMod(match=(value,mask), operation=DELETE\_STRICT, cookie) $
    		\EndIf

			\Return cmd
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[t]
    \caption{set Add abstraction}
    \label{alg:template}
    \begin{algorithmic}[1]
    \Require  id $i$.
%    \Ensure
    		\State $command \gets OFPGC\_ADD$
    		\State $type \gets OFPGT\_INDIRECT$
    		\State $group_id \gets i$
    		\State $buckets \gets [dummy\_backet]$
    		\State $cmd\gets GroupMod(command, type, group_id, buckets) $
			\Return cmd
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[t]
    \caption{set Delete abstraction}
    \label{alg:template}
    \begin{algorithmic}[1]
    \Require id $i$, unique constrain (defaults to False) $b\_unique$, caller id $caller$.
%    \Ensure
    		\State $command \gets OFPGC\_DELETE$
    		\State $group_id \gets i$
    		\State $cmd\gets GroupMod(command, group\_id) $
			\Return cmd
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[t]
    \caption{Update Algorithm with only a set}
    \label{alg:template}
    \begin{algorithmic}[1]
    \Require update requirements $U$, a set CUR.
    \Ensure installed policy is consistent with previous one
 		%\Do
 		\Repeat
 			\State read switch configuration
 			\State $pid\gets$ current policy id (only id in CUR)
 			\State $cur\_policy\gets$ current policy
 			\State $new\_policy\gets apply\_update(cur\_policy,U)$
 			\startTransaction
	 			\State CUR.Delete($pid$)
	 			\State CUR.Add($pid+1$)
	 			\State add $new\_policy$ rules %conditioned on $metadata=my\_a$
	 			%\State update policy selector to $my\_a$
 			\endTransaction
     		%\State $res \gets $ commit transaction
     	\Until{$res=True$}
     	%\doWhile{$res=False$}
%   \vspace{3mm}
			\Return

@@

In order to solve consensus we suggest to use the atomic multiple entries update capability (which is discussed in OpenFlow standard). A shared memory system that supports atomic write to (enough) multiple locations is known to allow the implementation of consensus objects. For example consider the following implementation utilizing $n^2$ memory locations, $M_{i,j}$, where $i,j\in[n]$, and the suggested values $\{v_i\}_{i\in [n]}$. All locations are initialized to zero.

@@

Our first observation is that a match-action configuration which is accessed by multiple controllers can be considered as shared memory. Note however that standard memory map and index/offset to a word, while Flow Table entries are not necessarily indexed. However, we assume that each flow entry has some property that allows the operator to specifically reference and update it, for example OpenFlow entry's cookie. In cases where this property value can't be determined by the operator, we can use other variable entry properties to store the index, for example we can represent memory value $v$ in offset $i$ by an match-action entry whose match is $in\_port = i$ and action is "forward port $x$".
%To overcome this we can use the match part of an entry as the index and the action part as the word value.

@@

The main idea of our solution is to adjust the previous policy update scheme in a way that performs consensus validation phase during every packet processing thereby requiring the servers only to make the first phase of the consensus - atomically writing the matrix.
In order to allow the packets to validate the consensus we make the following observation: while servers can be abstracted as memory writers, the packets processed by the switch are the readers. OpenFlow standard define an atomic write transaction as being atomic in relation to the processed packets thereby making packet processing a multiple read transaction but one read location per table is allowed. This means that if each memory location used in the consensus scheme will be stored in different table then each packet will be able to read all of them.

@@

@@

Our work is also closely related to the recent work on providing more high-level
abstractions and programming languages for SDNs, such as Frenetic~\cite{frenetic},
also enabling policy composition~\cite{pyretic}.

@@

This paper showed that powerful synchronization primitives
from distributed computing can be implemented in Openflow.
Our work complements existing research on the design of
distributed control planes, and also provides some missing links,
e.g., for the transactional approach taken by STN.
We also hope that our work can nourish the ongoing discussion of
what additional in-band features may be useful for future
versions of Openflow.

@@

>>>>

%\title{Solving Consensus with Openflow}

@@

%\title{One-Shot Coordination of Distributed SDN Controllers:\\On the Consensus Power of Openflow}

%\title{Coordinating Distributed SDN Controllers:\\On the Consensus Power of Openflow}

%\title{On the Synchronization Power of Openflow}

%\title{Synchronizing Distributed SDN Controllers\\{\Large The Openflow Container}}

\title{Synchronizing Distributed Openflow Controllers\\{\Large Toward More Powerful In-Band Synchronization Objects}}

@@

\begin{abstract}
Control planes of modern Software-Defined Networks (SDNs)
are \emph{distributed systems}: to ensure availability and fault-tolerance,
to improve load-balancing, and to reduce overheads,
control modules are physically distributed.
%a redundant control plane
%can ensure a high availability, and a spatially distributed
%control plane allows to handle time-critical dataplane
%events close to their origins.
%[[PK this is a contraversial point, not important for us
%Moreover, an SDN may be managed by multiple administrators simultaneously.
%]]
However, a distributed control plane also introduces
new challenges, and in order to guarantee a consistent network operation,
the actions of different controllers may need to be synchronized.
%[[PK sounds contradicting to the fact that such systems already exist
%However, not much is known today about how to implement
%such a distributed control plane.
%]]
In this paper, we introduce a novel synchronization object which 
we call \emph{container}.
The container can be implemented \emph{in-band}, using the standard Openflow 
protocol 
which support atomic multi-writes but only limited reads (namely 
testing the presence of rules). We describe how to  
leverage the container object to 
implement powerful consensus objects, transactions with all-or-nothing
 semantics, or atomically resolve
general conflicts between concurrent policy updates 
\emph{on-the-fly}, without affecting the ongoing traffic.
%[[PK
%In particular, we describe the implementation of test\&set primitives
%as well as multi-writer consensus objects which can be used
%to implement atomic transactions with all-or-nothing semantics.
%Interestingly, our approach allows controllers to detect and
%resolve very general notions of conflicts, not only, e.g., conflicts
%due to overlapping flow spaces but also conflicts based on network load.
%]]
\end{abstract}

@@

By consolidating and outsourcing the control over the dataplane switches to a logically
centralized controller, Software-Defined Networks (SDNs)
simplify network management and facilitate faster innovations:
the (software) control plane can evolve independently from the 
(hardware) data plane.
Moreover, Openflow, the standard SDN protocol today, introduces many generalizations,
in terms of traffic engineering, definition of flows, as well as in-band network functionalities,
by relying on a simple match-action paradigm which allows us to define
forwarding rules based not only on Layer-2, but also Layer-3 and Layer-4 (and possibly beyond).

While the logically centralized perspective offered by SDN is attractive,
there is a wide consensus that
the control plane should be physically \emph{distributed}.
First, in order to provide a high availability as well as a minimal degree of
fault-tolerance, controllers should be redundant~\cite{onix,stn,onos}: a failure
of one controller can be masked by other controllers. Second, it has also been proposed
to distribute controllers \emph{spatially}, in order to handle latency-sensitive and
communication intensive control plane events close to their origin.~\cite{devoflow,kandoo,jukka,disco}
Third, larger SDNs are likely to be operated by multiple administrators or may even offer
participatory interfaces where different users can install and trigger policy changes
concurrently~\cite{participatory,stn}.
All this, however, comes with the need for \emph{synchronizing}
the distributed controllers:
the network should be manipulated \emph{consistently}~\cite{cpc}.
This task is nontrivial~\cite{cap-theorem}.

@@

Consider, for example, the practical problem of
consistent installation of new forwarding policies, stipulating routes
that packets of different header spaces should follow across the
network.~\cite{network-update,roger-hotnets,correct,stn}
Installing conflicting forwarding rules, e.g., rules of the same priority defined over non-disjoint
flow spaces may lead to pathological network behavior (loops,
blackholes, etc.).
Similarly, installing diverging load-balancing policies may,
when combined, \emph{increase} the load~\cite{log-cent}.
To render things more difficult, controllers may also fail,
even before their updates have been completed.

%\textbf{}
\paragraph{Our Contributions}
In this paper, we show that
Openflow~\cite{of-spec}---the standard SDN protocol today---can
be used to implement powerful, high-level 
synchronization
objects \emph{in-band}.
In particular, we introduce a novel synchronization object 
called \emph{container}. A container can be seen as
a variant of the classic \texttt{test\&set} object
(known from distributed computing over shared memory),
which is tailored to Openflow: while Openflow supports 
atomic multi-write functionality (using bundling), 
read functionality is limited to testing the presence of
rules as well as flow space overlaps.

The container object is reminiscent of the hypothetical multi-write
shared-memory model which is known to enable solutions for important
synchronization problems, such as consensus, impossible to solve
otherwise~\cite{multi-objects}.
Indeed, we show how to exploit the container object to implement 
consensus among controllers. 
More generally, the container object turns out to be 
crucial for ensuring consistency of network
management on the control plane while preserving consistency of the
traffic processed on the data plane, and we show how to implement
\emph{packet-atomic} rule updates and more complex
\emph{atomic transactions}.
We also show how to replace the read-modify-write
primitives voluntarily assumed in the recent implementation of a
consistent composition system for forwarding policies~\cite{cpc}.

@@

\paragraph{Organization}
The rest of the paper is organized as follows.
In Section~\ref{sec:background}, we review the basics of the  Openflow
protocol and discuss the features which are relevant for our work.
We introduce our vision and the container consensus object in Section~\ref{sec:vision},
and present its implementation in Section~\ref{sec:container}.
We then discuss applications (Section~\ref{sec:apps}) as
well as extensions (Section~\ref{sec:discussion}).
After reviewing related literature in Section~\ref{sec:relwork}, we conclude
our contribution in Section~\ref{sec:conclusion}.

@@

\section{Openflow Background}\label{sec:background}

The synchronization mechanisms presented in this paper
are based on Openflow~\cite{of-spec}, the standard SDN protocol today. In this section,
we provide the necessary background.

In a nutshell,
Openflow
is based on a match-action concept: Openflow switches store
rules (installed by the controller) consisting of a match and an
action part. A packet matched by a certain rule will be subject
to the associated action (to be immediately applied now or later).
For example, an action can define a port to which the
matched packet should be forwarded, or add or change a tag
(a certain part in the packet header).
Each Openflow switch stores one or multiple flow tables,
each of which contains a set of rules (flow entries). Flow
tables form a pipeline, and flow entries are ordered according
to priorities: A packet arriving at a switch is first checked by
the rule of the highest priority in table 0: the fields of the
data packet are compared with the match fields of that rule,
and if they fit, some instructions (the actions) are executed;
subsequently, lower priority rules are checked. Depending on
the outcome of the table 0 processing, the packet may be sent
to additional flow tables in the pipeline; during the traversal of
the flow tables, the packet can be equipped with meta-tags to
transfer temporary information. Concretely, instructions can be
used to define additional tables to be visited (goto instruction),
to modify the set of to-be-applied actions (either by appending,
deleting, or modifying actions), or immediately apply some
actions to the packet. A meta-data field can be used to exchange
information between tables; it is part of the header that can be
inserted or removed from a packet via push and pop actions.
In general, a packet can be matched against any of its header
fields, and fields can be wildcarded and sometimes bitmasked.
(For instance, the meta-data field is maskable.) If no rule
matches, the packet is dropped.

The Openflow standard (as of \textbf{TODO}) offers three interesting
features which we will exploit to implement the transactional memory abstraction:

@@

\item \textbf{Conditional Application:}
The Openflow standard also supports the \emph{conditional} application of
certain types of control messages. A specific $\textsf{OFPFF\_CHECK\_OVERLAP}$ flag
set in the control message containing a new forwarding rule ensures
that the rule is applied if and only if it does not \emph{conflict}
with any other rule currently in the flow table.
This in-band conflict detection mechanism, combined with bundling,
allows us to implement specific types of \emph{transactions} that may
atomically install multiple rules if certain compatibility conditions
are met.

\item \textbf{Cookies:} Cookies allow to name and retrieve individual
flow rules. Our synchronization mechanism will implement controller-specific cookies. 
\end{enumerate}

\section{The Container Object}\label{sec:vision}

The container synchronization object introduced in this paper
is essentially a transactional memory abstraction which can be implemented
in standard Openflow switches. 
Before presenting the technical details in the next section,
we give a brief overview.

The postulated container object can store an arbitrary number of \emph{items}.
As we will see, these items correspond to flow entries on the Openflow switch.
The same item may even exist multiple times in the container (i.e., the container
is a multi-set).
The container
provides two main functions:
\begin{enumerate}
\item \textbf{\emph{Conditional Multi-Add($S$):}} A set $S$ of items $s\in S$
is (atomically) added to the container, if and only if no element $s\in S$
already existed in the container. If such an element already existed,
the operation does not take effect (i.e., the transaction is aborted).

\item \textbf{\emph{Conditional Multi-Remove($S$):}} A set $S$ of items $s\in S$
is (atomically) removed from the container, if and only if each element $s\in S$
already existed in the container. If one or more elements did not exist,
the operation does not take effect (i.e., the transaction is aborted).
\end{enumerate}

The container object allows controllers to atomically and
concurrently read and modify state. This  
allows us, for example, to solve consensus among any
number of controllers.
More generally, the container object
allows us to
achieve consistent network management.

\section{Openflow Implementation}\label{sec:container}

We now describe how to implement the container object
in Openflow. For our implementation, we will make
use of three main Openflow features: \emph{bundling},
\emph{conditional application}, and \emph{cookies}.
We will describe later how to implement some of these features
in older Openflow versions where they are not supported. \textbf{TODO}

A container contains items: they represent the flow table entries (i.e., a match-action rules, plus e.g., counters belonging to the rule).
Items can be added and removed atomically and in bundles (e.g., to update the switch with
a new policy). Recall that we want to add items only if they do not exist
yet; otherwise the transaction should abort without taking effect.
And similarly for removes: transactions should not take effect if the corresponding
update contains rules which currently do not exist at the switch.

Algorithm~\ref{alg:update} lies at the heart of our mechanism.
It takes a set of rules $U$ (not yet a policy), 
a multi-set $Next$ and a set (with full validation) $Cur$.
The algorithm proceeds as follows: first, we seek to
obtain a unique \emph{id}. FIXME: to be continued...

\begin{algorithm}[t]
    \caption{Advanced Update Algorithm}
    \label{alg:update}
    \begin{algorithmic}[1]
    \Require update requirements $U$, multi-set (with optional add validation) $Next$, set (with full validation) $Cur$.
    \Ensure installed policy is consistent with previous one
 		%\Do
 		\Repeat
 			\State \textbf{read} switch configuration
 			\State $my\_id\gets$ unused \emph{id} (in both $Cur$ and $Next$)
 			\State $pid\gets$ current policy id (only id in $Cur$)
 			\State $Next$\emph{.add}($pid$)
 			\State \textbf{read} switch configuration
 			\If {$pid\notin Cur$}
 				\State {\bf continue} (\textbf{restart} loop)
 			\EndIf
 			\State $cur\_policy\gets$ current policy
 			\State $new\_policy\gets apply\_update(cur\_policy,U)$
 			\startTransaction
 				\State $Next$\emph{.add}($my\_id$, True)
 				\State $Next$\emph{.delete}($my\_id$, True)
	 			\State DeleteGroup($pid$)
	 			\State AddGroup($my\_id$)
				\State $Next$\emph{.delete}($pid$)
	 			\State \textbf{add} $new\_policy$ rules %conditioned on $metadata=my\_a$
	 			%\State update policy selector to $my\_a$
 			\endTransaction
     		%\State $res \gets $ commit transaction
     	\Until{$res=\texttt{true}$}
     	%\doWhile{$res=False$}
%   \vspace{3mm}
			\Return

\begin{algorithm}[t]
    \caption{\textbf{\emph{Multi-Add()}}}
    \label{alg:multi-add}
    \begin{algorithmic}[1]
    \Require id $i$, unique constrain (defaults to \texttt{false}) $b\_unique$, caller id $caller$ (defaults to \emph{self}).
%    \Ensure
    		\State $value \gets i$
    		\State $mask \gets all-1$
    		\State $cookie \gets i.caller$
    		\State $action \gets i.caller$
    		\If {$b\_unique$}
    			\State $flag \gets 0$
    		\Else
    		    \State $flag \gets CHECK\_OVERLAP$
    		\EndIf
    					\State $cmd\gets \textsc{FlowMod}(match=(value,mask), operation = ADD, cookie, flag, action) $
			\Return $cmd$
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[t]
    \caption{\textbf{\emph{Multi-Remove()}}}
    \label{alg:template}
    \begin{algorithmic}[1]
    \Require id $i$, unique constrain (defaults to \texttt{false}) $b\_unique$, caller id $caller$.
%    \Ensure
    		\State $value \gets C.i$
    		\State $mask \gets all-1$
    		\State $cookie \gets C.i.caller$
    		\If {$b\_unique$}
    			\State $cmd\gets \textsc{FlowMod}(match=(value,mask), operation=DELETE) $
    		\Else
    		    \State $cmd\gets \textsc{FlowMod}(match=(value,mask), operation=DELETE\_STRICT, cookie) $
    		\EndIf

			\Return $cmd$
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[t]
    \caption{\textbf{\emph{Add()}}}
    \label{alg:template}
    \begin{algorithmic}[1]
    \Require  id $i$.
%    \Ensure
    		\State $command \gets OFPGC\_ADD$
    		\State $type \gets OFPGT\_INDIRECT$
    		\State $group_id \gets i$
    		\State $buckets \gets [dummy\_backet]$
    		\State $cmd\gets GroupMod(command, type, group_id, buckets) $
			\Return $cmd$
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[t]
    \caption{\textbf{\emph{Remove()}}}
    \label{alg:template}
    \begin{algorithmic}[1]
    \Require id $i$, unique constrain (defaults to False) $b\_unique$, caller id $caller$.
%    \Ensure
    		\State $command \gets OFPGC\_DELETE$
    		\State $group_id \gets i$
    		\State $cmd\gets GroupMod(command, group\_id) $
			\Return $cmd$
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[t]
    \caption{Update Algorithm with only a set}
    \label{alg:template}
    \begin{algorithmic}[1]
    \Require update requirements $U$, a set $Cur$.
    \Ensure installed policy is consistent with previous one
 		%\Do
 		\Repeat
 			\State \textbf{read} switch configuration
 			\State $pid\gets$ current policy id (only id in $Cur$)
 			\State $cur\_policy\gets$ current policy
 			\State $new\_policy\gets apply\_update(cur\_policy,U)$
 			\startTransaction
	 			\State $Cur$\emph{.delete}($pid$)
	 			\State $Cur$\emph{.add}($pid+1$)
	 			\State add $new\_policy$ rules %conditioned on $metadata=my\_a$
	 			%\State update policy selector to $my\_a$
 			\endTransaction
     		%\State $res \gets $ commit transaction
     	\Until{$res=\texttt{true}$}
     	%\doWhile{$res=False$}
%   \vspace{3mm}
			\Return

\section{Applications}\label{sec:apps}

FIXME: STN, consensus, ...

\paragraph{Consensus} FIXME: desribe how to implement consensus

\paragraph{One-Shot Update} FIXME: desribe how to implement one-shot update

\paragraph{STN} FIXME: desribe how to implement STN

\section{Discussion and Extensions}\label{sec:discussion}

FIXME: we could however even do stronger stuff: collision is only exact, not using masks: do not use the ternary nature

\paragraph{Beyond Container Objects}
FIXME: we can even do more general things!

\paragraph{Older Openflow Versions} FIXME: describe matrix approach

\section{In-Band Composition}\label{sec:compo}

FIXME: shall we?

\begin{comment}

\section{Transactions in Openflow}\label{sec:sync}

@@

@@

In order to solve consensus we suggest to use the atomic multiple entries update capability (which is discussed in Openflow standard). A shared memory system that supports atomic write to (enough) multiple locations is known to allow the implementation of consensus objects. For example consider the following implementation utilizing $n^2$ memory locations, $M_{i,j}$, where $i,j\in[n]$, and the suggested values $\{v_i\}_{i\in [n]}$. All locations are initialized to zero.

@@

Our first observation is that a match-action configuration which is accessed by multiple controllers can be considered as shared memory. Note however that standard memory map and index/offset to a word, while Flow Table entries are not necessarily indexed. However, we assume that each flow entry has some property that allows the operator to specifically reference and update it, for example Openflow entry's cookie. In cases where this property value can't be determined by the operator, we can use other variable entry properties to store the index, for example we can represent memory value $v$ in offset $i$ by an match-action entry whose match is $in\_port = i$ and action is "forward port $x$".
%To overcome this we can use the match part of an entry as the index and the action part as the word value.

@@

The main idea of our solution is to adjust the previous policy update scheme in a way that performs consensus validation phase during every packet processing thereby requiring the servers only to make the first phase of the consensus - atomically writing the matrix.
In order to allow the packets to validate the consensus we make the following observation: while servers can be abstracted as memory writers, the packets processed by the switch are the readers. Openflow standard define an atomic write transaction as being atomic in relation to the processed packets thereby making packet processing a multiple read transaction but one read location per table is allowed. This means that if each memory location used in the consensus scheme will be stored in different table then each packet will be able to read all of them.

@@

\end{comment}

@@

Our work is also closely related to the recent work on providing more high-level
abstractions and programming languages for SDNs, such as Frenetic~\cite{frenetic},
also enabling policy composition~\cite{pyretic}. Our work complements these works
on programming abstractoins
by initiating the discussion of concurrency objects and abstractions.

@@

This paper showed that powerful synchronization object
from distributed computing can be implemented in Openflow.
Our work complements existing research on the design of
distributed control planes, and also provides some missing links,
e.g., for the transactional approach taken by STN.
We also hope that our work can nourish the discussion of
which synchronization objects we can and want to implement
in an SDN. While much progress has been made over the
last years in the development of more intuitive programming
languages for SDNs (e.g., Frenetic, Pyretic, Merlin), the 
question of which synchronization abstractions to provide
has received much less attention.

@@

\begin{appendix}

\section{Backup Stuff}

using the bundling feature to implement \emph{multi-write} or, more
widely, \emph{multi-modify} transactions. .
, using a single Openflow switch that accepts
bundle control messages.

We first show that they can be used to replace the read-modify-write
primitives voluntarily assumed in the recent implementation of a
recent proposal for a consistent composition system for forwarding policies~\cite{cpc}.
We then use these transactions directly to implement a lightweight
forwarding policy composition system that requires no additional
synchronization on the control plane.
We then explore how more general classes of policies can be composed
\emph{on-the-switch}.

using the bundling feature to implement \emph{multi-write} or, more
widely, \emph{multi-modify} transactions. .
, using a single Openflow switch that accepts
bundle control messages.

We first show that they can be used to replace the read-modify-write
primitives voluntarily assumed in the recent implementation of a
recent proposal for a consistent composition system for forwarding policies~\cite{cpc}.
We then use these transactions directly to implement a lightweight
forwarding policy composition system that requires no additional
synchronization on the control plane.
We then explore how more general classes of policies can be composed
\emph{on-the-switch}.

we can do more than container:

we introduce container: it's not compare+set, ... its something new!
new object, is stronger than compare and set because bundle!

\end{appendix}
